Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-08-31 Manuscript.Rmd
much work on the methods, plus adding references from Mendeley through a bib file

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

@@

To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,564 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we took the exponential of the correlations, Fisher transformed them (i.e. tanh) so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately; 2) the concatenated timeseries from each pair of daily sessions; 3) each subject's dataset; and 4) on each step of a sliding window analysis (see Evaluation for more details).

@@

Where I is the identity matrix, D is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlations with every other vertex), and W is the correlation matrix. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler Vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected (Fiedler, 1975). Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (**Add to results**), spectral partitioning was unlikely to face the issues associated with its use in sparse networks (Fortunato & Hric, 2016).

@@

In order to evaluate the stability and topographical heterogeneity of the partitions within and across subjects, we chose the adjusted rand index (RI; Rand, 2012) as the reference metric of agreement between two clusterings. The adjusted rand index estimates the rate at which every pair of nodes in partition A were equally assigned to the same/different parcels on partition B, over all possible changes (e.g. clustered together on A, but separately on B). The resulting ratio is compared to a baseline given by the expectation based on a random vertex assignment for an equal number of clusters across partitions. Because of this, the adjusted rand index is a simple metric that determines the percentage agreement of any two partitions that is agnostic to the labeling scheme. *(SHOULD WE INCLUDE THIS?) On the other hand, variation of information gives a general sense of the amount of information to be gained by the complementary community partitioning (Meila, 2007), measured by the addition of the conditional entropies per combination of clusters across partitions. While the rand index is more easily interpreted, it suffers from the assumption that the baseline should have an equal number of vertices per cluster across partitions, which is not necessarily the case. Therefore, I used variation of information as a more robust similarity metric (Fortunato & Hric, 2016).*

>>>>

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  word_document: default
  pdf_document: default
  html_document: default
bibliography: mPFC_paper.bib
---

@@

To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,564 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and took the exponential of the correlations so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately; 2) the concatenated timeseries from each pair of daily sessions; 3) each subject's dataset; and 4) on each step of a sliding window analysis (see Evaluation for more details).

@@

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlations with every other vertex), and $W$ is the correlation matrix. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler Vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected (Fiedler, 1975). Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (**Add to results**), spectral partitioning was unlikely to face the issues associated with its use in sparse networks (Fortunato & Hric, 2016).

@@

In order to evaluate the stability and topographical heterogeneity of the partitions within and across subjects, we chose the adjusted rand index (RI; Rand, 2012) as the reference metric of agreement between two clusterings. The adjusted rand index estimates the rate at which every pair of nodes in partition A were equally assigned to the same/different parcels on partition B, over all possible changes (e.g. clustered together on A, but separately on B). The resulting ratio is compared to a baseline given by the expectation based on a random vertex assignment for an equal number of clusters across partitions. Because of this, the adjusted rand index is a simple metric that determines the percentage of agreement between any two partitions that is agnostic to the labeling scheme. *(SHOULD WE INCLUDE THIS?) On the other hand, variation of information gives a general sense of the amount of information to be gained by the complementary community partitioning (Meila, 2007), measured by the addition of the conditional entropies per combination of clusters across partitions. While the rand index is more easily interpreted, it suffers from the assumption that the baseline should have an equal number of vertices per cluster across partitions, which is not necessarily the case. Therefore, I used variation of information as a more robust similarity metric (Fortunato & Hric, 2016).*
